### 2022-02-22 writing/ACADIA21/essay_latex/examples via tomas/chapter_TNA/TNA.tex

\chapter{Thrust Network Analysis}
\label{chap:tna}

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor
incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

Figure \ref{fig:texas_model} in page \pageref{fig:texas_model} is cool and i
have nothing else to say about it. Do you like it?

\begin{equation}
D_{50} =   \frac{\displaystyle{\int^{50}_{0}p^{2}(t)
\mathrm{d}t}}{\displaystyle{\int^{\infty}_{0}p^{2}(t) \mathrm{d}t}}   \quad \mathrm{(dB)}
\end{equation}

\section{Some Stuff}
\label{sec:some_stuff}

Deb \cite{Deb2001} talks about GA's and blah blah blah\ldots. He probably said
something like this:

\begin{quotation}
``Time and motion studies show that the fewer kinds of materials there are, the more efficient factory assembly is - and therefore demand a certain simplicity in the variety of materials used. This need for simplicity conflicts with the fact that the form will function better if we choose the best material for each separate purpose separately. But then, on the other hand, functional diversity of materials makes for expensive and complicated joints between components, which is liable to make maintenance less easy. Further still, all three issues, simplicity, performance and jointing, are at odds with our  to minimize the cost of materials\cite{tms2013c}. For if we choose the cheapest material for each separate task, we shall not necessarily have simplicity, nor optimum performance, nor materials which can be cleanly jointed.''
\end{quotation}

 Lorem ipsum dolor  sit amet, consectetur adipiscing elit, sed do
 eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

  Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor
 incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

\begin{figure}[ht]%[htbp]
\begin{center}
\begin{center}\includegraphics[width = 1\textwidth]{texas_model.jpg}\end{center}
\caption{TNA dual graph of the MLK memorial blah blah\ldots}
\label{fig:texas_model}
\end{center}
\end{figure}

  Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor
 incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

  Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor
 incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

  Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor
 incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

 This is some pseudocode example from an awesome PhD thesis by Tom\'{a}s
 M\'{e}ndez Echenagucia:\\

 \begin{algorithmic}[1]
\State Generate a random and coded population 
\For{$i\gets 1$, number of Generations}
    \State Decode and Scale the population
    \State Calculate fitness values for the population
    \State Run Selection operation
    \State Run Crossover operation
    \State Run Mutation operation
    \State Perform Exit condition test

    \If{Exit condition test = \texttt{True}}
        \State \texttt{Exit} GA
    \ElsIf{Exit condition test = \texttt{False}}
        \State Continue to the next Generation
    \EndIf
\EndFor
\end{algorithmic}

I told you it was awesome.